\propertitle

### 2019-12-07 maxNt_v6.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-12-07T15:36:17+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Inferring the total activity of a large neuronal population\\ from a small sample}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

% Before applying the formula above to two concrete data sets we want to add
% two remarks about the maximum-entropy method that are seldom made in the
% neuroscientific literature. They are important for the interpretation of
% the results and are further discussed in \sect~\ref{sec:assumptions}.
% First, the maximum-entropy
% method % (based on the \emph{Shannon} relative
% % entropy, as opposed to other entropy functions such as Burg's
% % \citeyear{burg1975})
% rests on some implicit assumptions about the probabilities for the long-run
% frequency distribution of activities, besides the assumptions just
% mentioned at the end of the previous paragraph. So the often-heard
% statement that it gives \enquote{the maximally unbiased (or non-committal)
%   distribution} must be taken with a grain of salt.
% \citep{jaynes1986d_r1996,portamana2009,portamana2017}. Second,
A maximum-entropy distribution like $\px(\yA \| M, N)$ is equivalent to the
zeroth-order approximation (in the sense of Laplace's approximation
\citep[\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000}) of
four distinct distributions for the larger population, which differ
numerically from one another in higher-order approximations:
\begin{enumerate}[label=(\roman*)]
\item\label{item:prob_freq} the most probable \emph{frequency} distribution
  for the total activity across the \emph{recorded} bins, %$\yFF(\yA)$,
\item the \emph{belief} distribution for the value of the total
  activity at any time bin among the \emph{recorded} ones,
  %$\yp(\yA_{t} \| \yH)$ with $1\le t \le T$,
\item the most probable \emph{frequency} distribution for the total
  activity in a very long run of \emph{new} time bins, %$\ynuu(\yA)$,
\item the \emph{belief} distribution for the value of the total
  activity at a \emph{new} time bin.
  % $\yp(\yA_{t} \| \yH)$ with $t<1$ or $t>T$
\end{enumerate}
Their mathematical relationship is presented more explicitly in
appendix~\ref{sec:inference_prob}. The
distribution~\eqref{eq:app_maxent_pop_first} can therefore be interpreted
in each of the four ways above, in this approximate sense. In the present
case the validity of the approximation decreases as the ratio $nN/T$
increases. The most robust interpretation is~\ref{item:prob_freq}; and this
is how we prefer to interpret the distribution $\px$: as \emph{the most
  probable relative-frequency distribution} of the activity $\yA$. See
\sect~\ref{sec:assumptions} for further discussion about assumptions and
approximations.

@@

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-12-07T17:52:23+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Inferring the total activity of a large neuronal population\\ from a small sample}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

% Before applying the formula above to two concrete data sets we want to add
% two remarks about the maximum-entropy method that are seldom made in the
% neuroscientific literature. They are important for the interpretation of
% the results and are further discussed in \sect~\ref{sec:assumptions}.
% First, the maximum-entropy
% method % (based on the \emph{Shannon} relative
% % entropy, as opposed to other entropy functions such as Burg's
% % \citeyear{burg1975})
% rests on some implicit assumptions about the probabilities for the long-run
% frequency distribution of activities, besides the assumptions just
% mentioned at the end of the previous paragraph. So the often-heard
% statement that it gives \enquote{the maximally unbiased (or non-committal)
%   distribution} must be taken with a grain of salt.
% \citep{jaynes1986d_r1996,portamana2009,portamana2017}. Second,
A maximum-entropy distribution like $\px(\yA \| M, N)$ is equivalent to the
zeroth-order approximation (in the sense of Laplace's approximation
\citep[\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000}) of
four distinct distributions for the larger population, which differ
numerically from one another in higher-order approximations:
\begin{enumerate}[label=(\roman*)]
\item\label{item:prob_freq} the most probable \emph{frequency} distribution
  for the total activity across the \emph{recorded} bins, %$\yFF(\yA)$,
\item\label{item:prob_binin} the \emph{belief} distribution for the value
  of the total activity at any time bin among the \emph{recorded} ones,
  %$\yp(\yA_{t} \| \yH)$ with $1\le t \le T$,
\item\label{item:prob_longrun} the most probable \emph{frequency}
  distribution for the total activity in a very long run of \emph{new} time
  bins, %$\ynuu(\yA)$,
\item\label{item:prob_binout} the \emph{belief} distribution for the value
  of the total activity at a \emph{new} time bin.
  % $\yp(\yA_{t} \| \yH)$ with $t<1$ or $t>T$
\end{enumerate}
Their mathematical relationship is presented more explicitly in
appendix~\ref{sec:inference_prob}. The
distribution~\eqref{eq:app_maxent_pop_first} can therefore be interpreted
in each of the four ways above, in this approximate sense. In the present
case the validity of the approximation decreases as the ratio $nN/T$
increases. The most robust interpretation is~\ref{item:prob_freq}; and this
is how we prefer to interpret the distribution $\px$: as \emph{the most
  probable relative-frequency distribution} of the activity $\yA$. See
\sect~\ref{sec:assumptions} for further discussion about assumptions and
approximations.

@@

From the formulae above we can also find the four distributions mentioned
at the end of \sect~\ref{sec:model}. The frequency
distribution~\ref{item:prob_freq} is the mode of $\pf(\yF \| \yf,\yI)$. The
belief distribution~\ref{item:prob_binin} is
\begin{equation}
  \label{eq:prob_binin}
  \pf(\yA_{t} \| \yf, \yI) =
  \sum_{\yF} \yFF_{\yA_{t}}\; \pf(\yF \| \yf, \yI),
  \qquad t \in \set{1,\dotsc,T}.
\end{equation}
The frequency distribution~\ref{item:prob_longrun} is the mode of $\pf(\ynu
\| \yf,\yI)$. Finally, the belief distribution~\ref{item:prob_binout} is
\begin{equation}
  \label{eq:prob_binout}
  \pf(\yA_{t} \| \yf, \yI) =
  \int\!\!\!\di\ynu\;\nu_{\yA_{t}}\;\pf(\ynu \| \yf, \yI),
  \qquad t \notin \set{1,\dotsc,T}.
\end{equation}
